---
title: "Rasch analysis with Bayesian and frequentist methods"
subtitle: "A comparison using NHANES data for PHQ-9"
author:
  name: 'Magnus Johansson'
  affiliation: 'Karolinska Institutet'
  orcid: '0000-0003-1669-592X'
  email: 'magnus.johansson.3@ki.se'
  degrees: PhD
  url: https://ki.se/personer/magnus-johansson-3
date: 2026-03-31
date-format: iso
date-modified: last-modified
google-scholar: true
toc: true
number-sections: true
syntax-highlighting: github
toc-depth: 4
citation:
  type: 'webpage'
format: 
  html:
    code-fold: false
execute: 
  cache: true
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
bibliography: rasch_bayes_comp.bib
---

```{r}
library(easyRasch) # only available on GitHub, "pgmj/easyRasch"
library(iarm)
library(eRm)
library(mirt)
library(tidyverse)
library(brms)
library(loo)
options(mc.cores = 8)
set.seed(92756246)
library(priorsense)
library(tidybayes)
library(easyRaschBayes) # dev version 0.2.0.1
library(haven)

select <- dplyr::select
count <- dplyr::count
rename <- dplyr::rename
```

## Introduction

This is an example Rasch analysis using frequentist methods and their Bayesian counterparts, as implemented in the `easyRaschBayes` R package. The frequentist methods are used through the `easyRasch` package, which is mostly a wrapper package except for some [bootstrap/simulation functions](https://pgmj.github.io/simcutoffs.html) [see also @johansson_detecting_2025]. As such, the sections above will highlight which R package contains the frequentist function. For a more extensive discussion on the metrics and their interpretation, see the [`easyRasch` vignette](https://pgmj.github.io/raschrvignette/RaschRvign.html).

## Importing data

Data files manually downloaded from [September 2024 NHANES data website](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Questionnaire&CycleBeginYear=2024). As the dataset is quite large with 6337 observations, we will select a subsample of n = 750 with complete responses to all 9 items of the PHQ-9 and a reasonable amount (10%) of respondents with all zeroes.

```{r}
#| eval: false
d_all <- read_xpt("data/DPQ_L_2024_nhanes.xpt")

d_sub <- d_all[2:10] %>% 
  mutate(across(everything(), ~ car::recode(.x, "4:9=NA"))) %>% #only responses 0-3 are valid
  na.omit() %>% 
  set_names(paste0("q",1:9)) %>% 
  mutate(sumscore = rowSums(.))

d_sub %>% 
  summarise(mean(sumscore == 0)) # 24.3%

d_zero <- d_sub %>% 
  filter(sumscore == 0) %>% 
  slice_sample(n = 75)

d <- d_sub %>% 
  filter(sumscore > 0) %>% 
  slice_sample(n = 675) %>% 
  bind_rows(d_zero) %>% 
  select(!sumscore)

saveRDS(d, "data/phq9_nhanes2024_subsample.rds")
```

```{r}
d <- readRDS("data/phq9_nhanes2024_subsample.rds")
head(d)
```

All PHQ-9 questions/items are introduced with the phrase

> Over the last 2 weeks, how often have you been bothered by the following problems

All items have the same ordinal response categories:

- "Not at all"
- "Several days"
- "More than half the days"
- "Nearly every day"

We'll set up vectors for item and response category labels for later use in plots, etc.

```{r}
item_desc <- c(
  "Little interest or pleasure in doing things",
  "Feeling down, depressed, or hopeless",
  "Trouble falling or staying asleep, or sleeping too much",
  "Feeling tired or having little energy", "Poor appetite or overeating",
  "Feeling bad about yourself - or that you are a failure or have let yourself or your family down",
  "Trouble concentrating on things, such as reading the newspaper or watching television",
  "Moving or speaking so slowly that other people could have noticed?",
  "Thoughts that you would be better off dead or of hurting yourself in some way"
)

item_resp <- c("Not at all","Several days","More than \nhalf the days","Nearly every day")
```


## Response distributions

Using functions from `easyRaschBayes` with the wide format dataset.

::: panel-tabset
### Bars
```{r}
#| fig-height: 8
plot_bars(d, text_y = 10, category_labels = item_resp)
```

### Stacked bars
```{r}
#| fig-height: 8
plot_stackedbars(d, text_size = 4.5, category_labels = item_resp) + 
  theme_minimal(base_size = 14)
```

### Tile plot
```{r}
#| fig-height: 7
plot_tile(d, category_labels = item_resp, 
          item_labels = str_wrap(item_desc,24)) + 
  theme_minimal(base_size = 13)
```
:::

Low frequency use of upper categories of item 9, which is expected in this normal population sample, even if we manipulated the sampling a bit to minimize the floor effects.

## Fit `brms` model

For `brms`, we need data in long format and with the lowest category coded as 1.

```{r}
dl <- d %>% 
  mutate(across(everything(), ~ .x + 1)) %>% 
  rownames_to_column("id") %>% 
  pivot_longer(!id, 
               names_to = "item", 
               values_to = "response")
```

Now we can fit a Rasch partial credit model (PCM) with a weakly informative prior.

```{r}
#| eval: false

prior_pcm <- prior(normal(0, 3), class = "Intercept") +
  prior(normal(0, 3), class = "sd", group = "id")

fit <- brm( 
  response | thres(gr = item) ~ 1 + (1 | id),
  data   = dl,
  family = acat("logit"),
  prior = prior_pcm,
  chains = 4,
  cores  = 4, 
  iter   = 4000,
  warmup = 1500,
  save_pars = save_pars(all = TRUE)
)

saveRDS(fit,"data/phq9fit.rds")
```

The model took about 175s to fit on my Macbook Pro with M4 Max chip.

```{r}
fit <- readRDS("data/phq9fit.rds")
```

## `brms` model fit

First, we'll run some basic diagnostics for the model.

```{r}
summary(fit) # review Rhat and ESS metrics
fit <- add_criterion(fit, criterion = "loo")
powerscale_sensitivity(fit, get_variables(fit)[1:28])
```

There are additional posterior predictive checks that could be run as well, using `pp_check()`. [A recent preprint](https://www.journalovi.org/2025-sailynoja-visual-predictive-checks/#sec-visual-predictive-checks-for-ordinal-data) presents some new types of pp-checks that seem highly relevant but AFAIK has not been implemented in the `bayesplot` package yet. The same paper also critiques the usefulness of `pp_check(type = "bars_grouped")` [@sailynoja_recommendations_2025].

## Rasch analysis

For this example, I have opted to use 2000 draws available from the `brmsfit` object. It takes quite a bit of time to extract and compute a large number of draws for the various functions. I would suggest to use 1000 draws for a casual first analysis, and probably 4000+ for a publication ready analysis. This is just based on small scale experimenting, so feel free to run a simulation study if you want hard evidence, or rely on Kruschke's recommendation to have an ESS (effective sample size) of at least 10 000 draws [@kruschke_rejecting_2018] to be on the safe side.

### Conditional item infit

See Müller [-@muller_item_2020] and Johansson [-@johansson_detecting_2025] for more details on this metric and particularly Müller on why the traditional unconditional item fit should not be used with sample sizes above 200. The `easyRasch` package uses parametric bootstrap to determine adaptive cutoff values for every item.

The Bayesian model doesn't need bootstrap, since we can get expected and observed values directly from the model fit object.

::: panel-tabset
#### iarm/eRm
```{r}
simfit <- RIgetfit(d, iterations = 400, cpu = 8)
RIitemfit(d, simfit)
RIgetfitPlot(simfit, d)
```

Item 2 is strongly overfit.
- 2: "Feeling down, depressed, or hopeless"

Item 3 is strongly underfit, and item 5 is borderline underfit.
- 3: "Trouble falling or staying asleep, or sleeping too much"
- 5: "Poor appetite or overeating"

The two psychosomatic symptoms (item 3 and 5) deviate from other items, likely due to multidimensionality, while item 2 is a very general statement that is strongly related to the latent construct.

#### Bayes
```{r}
inf1 <- infit_statistic(fit, ndraws_use = 2000)
inf1post <- infit_post(inf1)
inf1post$summary
inf1post$hdi
inf1post$plot
```

Based on ppp-values:

- item 2 is overfit
- item 3, 5, and 8 are underfit
  - 8: "Moving or speaking so slowly that other people could have noticed?"
  
:::

### Item-restscore

Using Goodman-Kruskal's gamma [@kreiner_note_2011].

::: panel-tabset
#### iarm
```{r}
RIrestscore(d)
```

- Items 2, 6 and 9 are overfit
- Item 3 is underfit.
  - 3: "Trouble falling or staying asleep, or sleeping too much"
  
#### Bayes
```{r}
irest1 <- item_restscore_statistic(fit, ndraws_use = 2000)
irest1post <- item_restscore_post(irest1)
irest1post$summary
irest1post$plot
```

Since the observed item-restscore is non-parametric, there is no posterior
distribution, only a point estimate.

Based on ppp-values:

- items 2, 6 and 9 are overfit
- items 3 and 5 underfit
  - 3: "Trouble falling or staying asleep, or sleeping too much"
  - 5: "Poor appetite or overeating"
:::

### Posterior Predictive Item Fit

Based on code in Bürkner [-@burknerAnalysingStandardProgressive2020] and adapted for polytomous data.

```{r}
# Categorical log-likelihood criterion (for polytomous models)
ll_categorical <- function(y, p) log(p)

ll_fit <- fit_statistic_pcm(fit, 
                  criterion = ll_categorical, 
                  ndraws_use = 1000,
                  group = item)

# Summarise: posterior predictive p-values per item
ll_fit %>%
  group_by(item) %>%
  summarise(
    observed   = mean(crit),
    replicated = mean(crit_rep),
    ppp        = mean(crit_rep > crit)
  )

# Use ggplot2 to make a histogram
ll_fit %>%
  ggplot(aes(crit_diff)) +
  geom_histogram(aes(fill = ifelse(crit_diff > 0, "above","below"))) +
  facet_wrap("item") +
  theme_bw() +
  theme(legend.position = "none")
```

Based on ppp:

- item 2 is overfit
- item 3 is underfit, maybe also item 5

### Conditional ICC

::: panel-tabset
#### iarm
```{r}
RIciccPlot(d)
```

#### Bayes
```{r}
plot_icc(fit)

```
The theta values are off in the 0.2.0 CRAN version, this is fixed in the GitHub development version 0.2.0.1.
:::


### Local dependency

Using Yen's Q3 statistic [@christensen_critical_2017], again with parametric bootstrap to determine a critical value.

::: panel-tabset
#### mirt
```{r}
simcor <- RIgetResidCor(d, iterations = 1000, cpu = 8)
RIresidcorr(d, simcor$p99)
```

Based on 1000 simulations and the 99th percentile as a cutoff.

Correlated item pairs, sorted on strongest correlations

- 1 & 2
  - 1: "Little interest or pleasure in doing things"
  - 2: "Feeling down, depressed, or hopeless"
- 2 & 6
  - 6: "Feeling bad about yourself - or that you are a failure or have let yourself or your family down"
- 6 & 9
  - 9: "Thoughts that you would be better off dead or of hurting yourself in some way"

#### Bayes
```{r}
q3_1 <- q3_statistic(fit, ndraws_use = 2000)
q3_1post <- q3_post(q3_1, n_pairs = 8)
q3_1post$summary %>% head(8)
q3_1post$hdi %>% head(8)
```

Item pairs most strongly correlated:

- 2 & 6
- 1 & 2
- 6 & 9

A bit less strong but still likely too much:

- 2 & 9
- 1 & 6

These seem to be clustered together.

```{r}
#| fig-height: 7
q3_1post$plot
```
:::


### PCA of residuals

This is primarily a way to evaluate patterns in the standardized model residuals.

::: panel-tabset
#### Bayes
```{r}
#| cache.lazy: false
pca1 <- plot_residual_pca(fit, ndraws_use = 2000)
pca1
```

#### eRm/stats
```{r}
RIpcmPCA(d, rotation = "none")
simpca <- RIbootPCA(d, iterations = 500, cpu = 8)
simpca$p95
simpca$p99

RIloadLoc(d) +
  coord_flip()
```

:::

### Response category probabilities

A.k.a. Item Probability Functions

::: panel-tabset
#### mirt
```{r}
mirt(d, model=1, itemtype='Rasch', verbose = FALSE) %>% 
  plot(type="trace", as.table = TRUE, 
       theta_lim = c(-3,5))
```
#### Bayes
```{r}
plot_ipf(fit, theta_range = c(-3,5),
         category_labels = item_resp,
         ribbon_alpha = 0.25)
```
:::

### Targeting

::: panel-tabset
#### eRm
```{r}
RItargeting(d, fast_thetas = T, 
            xlim = c(-5,3))
```
#### Bayes
```{r}
plot_targeting(fit)
```
:::

### DIF

For the case of demonstration, we'll generate a random "gender" variable.

```{r}
#| eval: false
d2 <- d %>% 
  mutate(gender = factor(sample(c("M", "F"), n(), TRUE)))

dl2 <- d2 %>% 
  mutate(across(where(is.numeric), ~ .x + 1)) %>% 
  rownames_to_column("id") %>% 
  pivot_longer(!c(id,gender), 
               names_to = "item", 
               values_to = "response")
saveRDS(list(d2,dl2), file = "data/phq9difdata.rds")
```

```{r}
difdata <- readRDS("data/phq9difdata.rds")
d2 <- difdata[[1]]
dl2 <- difdata[[2]]
```


First, we need to check the response distribution for both gender categories.

```{r}
RIdifTileplot(d, d2$gender)
```

No empty cells, which is good, particularly in the lower categories.

::: panel-tabset
#### eRm table
```{r}
RIdifTableLR(d, dif.var = d2$gender)
```
#### eRm figure
```{r}
RIdifFigureLR(d, dif.var = d2$gender)
```
#### eRm threshold figure
```{r}
RIdifThreshFigLR(d, dif.var = d2$gender)
```
#### iarm CICC
```{r}
RIciccPlot(d, dif = "yes", dif_var = d2$gender)
```

#### Bayes CICC

Does not require refitting the model.

```{r}
plot_icc(fit, 
         dif_var = "gender", 
         dif_data = dl2)
```
#### Bayes DIF

This does require refitting the model with the DIF variable included, which will take quite a bit of time. I'll preload the one below, since it took almost 30 minutes on 4 cores.

```{r}
#| eval: false
phq9_dif_gender <- dif_statistic(fit, group_var = "gender", data = dl2)
saveRDS(phq9_dif_gender, "data/phq9_dif_gender_fit.rds")
```

You can also use `dif_statistic()` to just output a formula for `brms` and refit the model manually.

```{r}
dif_statistic(fit, group_var = "gender", data = dl2, refit = FALSE)
```

For polytomous models, it is often also of interest to review the potential effect of DIF on individual threshold locations, which can be done with `dif_type = "non-uniform"`.

```{r}
phq9_dif_gender <- readRDS("data/phq9_dif_gender_fit.rds")

phq9_dif_gender$summary %>% 
  knitr::kable()
phq9_dif_gender$plot
```
:::

## Summary

These PHQ-9 data clearly show several significant psychometric weaknesses. Several items show strong misfit to the measurement model, residual correlations, and disordered item category threshold locations. A next step in the analysis process would be to remove one or two misfitting items, in this data items 3 (underfit) and 2 (overfit and having residual correlations), and run a new model. Since this is  intended primarily as a demonstration of the functions available in `easyRaschBayes`, I'll leave the continued analysis to the curious reader.

However, there are some functions left to demonstrate that would normally only be used when having arrived at a set of items with adequate psychometric properties. These include reliability and item and person parameters. 

For this last part, we'll use a model including only PHQ items 1, 4, 6, and 7, since these make up a well-functioning item set regarding item fit and LD. Item threshold locations are not optimally separated, as will be shown after we look at the recently proposed reliability metric Relative Measurement Uncertainty [RMU, @bignardi_general_2025].

### Reliability

First, we refit the model, modified to include only four items.

```{r}
#| eval: false
dl2 <- dl %>% 
  filter(item %in% paste0("q",c(1,4,6,7)))
fit_4item <- update(fit, newdata = dl2)
saveRDS(fit_4item,"data/phq9fit4items.rds")
```

Next, the draws of person parameters are extracted from the model.

```{r}
fit_4item <- readRDS("data/phq9fit4items.rds")

posterior_draws = fit_4item %>%
  as_draws_df() %>%
  as_tibble() %>% 
  select(starts_with("r_id")) %>%
  t()

RMUreliability(posterior_draws)
```

### Item category probabilities

```{r}
plot_ipf(fit_4item)
itemps <- item_parameters(fit_4item)
itemps$threshold_order
```

::: panel-tabset
### Item locations (brms)

```{r}
itemps$locations_wide %>% 
  arrange(item) %>% 
  kable()
```

### Item locations (eRm)

```{r}
RIitemparams(d[,c(1,4,6,7)])
```
:::

### Person locations

We get EAP, expected a-posteriori, from `brms`, but Weighted Likelihood Estimation [WLE, @warm_weighted_1989] is also available in the function output. WLE has been shown to have slightly less bias [see also @kreiner_specific_2025]. Note that the extreme values cannot be estimated using WLE, but are set at the `theta_range` boundaries. Standard errors of measurement is also included in the output object.

```{r}
personps <- person_parameters(fit_4item, theta_range = c(-4,4))
hist(personps$person_estimates$eap, breaks = 20)
hist(personps$person_estimates$wle, breaks = 20)
cor.test(personps$person_estimates$eap,
         personps$person_estimates$wle)
```

For comparison, here are the WLE person parameters from `eRm`.

```{r}
eRm_ppar <- RIestThetas(d[,c(1,4,6,7)])
hist(eRm_ppar$WLE, breaks = 20)
```

### Transformation table

When data fits the Rasch model, the ordinal sum score is a sufficient statistic for the latent variable. However, the ordinal sum score is not an interval scale variable and it is likely beneficial to transform the sum score to EAP or WLE prior to further statistical analysis.

::: panel-tabset
#### Bayes
```{r}
personps$score_table %>% 
  kable()
```
#### eRm/WLE
```{r}
RIscoreSE(d[,c(1,4,6,7)], output = "table")
```
:::

## References