---
title: "Data retrieval with R using API calls"
subtitle: "The Swedish National Agency for Education database API"
author:
  - name: 'Magnus Johansson'
    affiliation: 'RISE Research Institutes of Sweden'
    affiliation-url: 'https://ri.se/shic'
    orcid: '0000-0003-1669-592X'
  - name: 'Jens Mattsson'
    affiliation: 'RISE Research Institutes of Sweden'
    affiliation-url: 'https://ri.se/shic'
date: last-modified
date-format: iso
execute: 
  cache: true
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

## Background

This is a post documenting explorative work to retrieve data using the [database API](https://www.skolverket.se/om-oss/oppna-data/api-for-skolenhetsregistret) maintained Swedish National Agency for Education. The conditions for using the API and database are described in the [CC0 1.0 license](https://creativecommons.org/publicdomain/zero/1.0/deed.en). The API is documented here: <https://api.skolverket.se/skolenhetsregistret/swagger-ui/index.html>.

::: {.callout-note}
### Note
Please note that this is just documentation of our work, not a guide. There will be mistakes and suboptimal routes taken. But in the end I hope we will produce something that may be useful to others. Not everyone will be interested in using the API for the same purpose as we had, so hopefully our troubleshooting will make the potential use wider.
:::

A lot of the output will be in Swedish, but you will probably be able to follow along even if your Swedish knowledge is limited. Basic word list:

- school = skola
- school unit code = Skolenhetskod

The purpose of this is two-fold. First, various kinds of data on school and municipality levels are of interest in the project ["Data in dialogue"](https://www.ri.se/en/what-we-do/projects/data-in-dialogue-risk-and-protective-factors-for-children-and-youth). Second, in order to conduct analysis of missing data and selection bias we need demographic data about students at the schools participating in school surveys that are used to assess risk and protective factors.

::: {.callout-note}
### Note
While this blog post is written by me (Magnus), a lot of the initial trial and error work was done by my colleague Jens Mattsson.
:::

## Setting up

```{r}
library(httr)
library(arrow)
library(tidyverse)
library(rjson)
library(jsonlite)
library(glue)
library(readxl)
```

First, let's get a list of municipalities and their codes. The first two numbers in the four number code denotes the region. Sweden has 290 municipalities and 21 regions. For this project, we are interested in the regions of Stockholm and Uppsala, which have codes 01 and 03.

```{r}
municipalities <- read_parquet("2023-03-28_KOLADA_Municipality_list.parquet") %>% 
  filter(str_detect(id, "^01|^03")) %>% 
  select(!type)

glimpse(municipalities)
```

## Getting data for one municipality

Looking at the specifications of the API, we should be able to get all schools in a municipality by doing a call according to `/v1/kommun/{municipalityCode}`. Unfortunately, the API does not seem to allow making one call for multiple municipalities. The base URL is `https://api.skolverket.se/skolenhetsregistret`.

We'll start by getting data from one municipality.

```{r}
data <- GET("https://api.skolverket.se/skolenhetsregistret/v1/kommun/0127") %>% 
  content("text") %>% 
  fromJSON()

glimpse(data)
```

We get a list of 3, where the data of interest seems to be within the nested dataframe `$Skolenheter`. Let's subset that into its own dataframe object.

```{r}
schools <- data$Skolenheter
glimpse(schools)
```

Now, this is just a list of schools and their unit codes (`schools$Skolenhetskod`), it contains no data. But we need this list to know which school unit codes to retrieve data for. There is also a `Status` variable which seems to have the options of active or not.

```{r}
schools %>% 
  count(Status)
```

There was also a third option for "planned". Not sure how to use this information at this point.

### Data from one school

We'll retrieve data for one school first.

```{r}
sdata <- GET("https://api.skolverket.se/skolenhetsregistret/v1/skolenhet/84411355") %>% 
  content("text") %>% 
  fromJSON()

glimpse(sdata)
```

This provides a lot of information about the school itself, which can be useful. There is also a version 3 of the API, which contains more information:

```{r}
sdataV3 <- GET("https://api.skolverket.se/planned-educations/v3/school-units/84411355") %>% 
  content("text") %>% 
  fromJSON()

glimpse(sdataV3)
```

Can we find statistics for the school?

```{r}
test <- GET("https://api.skolverket.se/planned-educations/v3/school-units/84411355/statistics") %>% 
  content("text") %>% 
  fromJSON()

glimpse(test)
```

No data, but some clues:

```{r}
test$body$`_links`$`gr-statistics`
```

We'll try that URL.

```{r}
test <- GET("https://api.skolverket.se/planned-educations/v3/school-units/84411355/statistics/gr") %>% 
  content("text") %>% 
  fromJSON()

glimpse(test)
```

It seems like we need to specify the type of school to retrieve the stats. In this case, "gr" for "grundskola", which corresponds to classes 1-9 in Sweden (ages ~ 7-15).

### Data wrangling

Some data wrangling will be needed to get the list() format data into a dataframe that can be used as a template for downloading and merging data for all schools we are interested in.

```{r}
stats <- test$body %>% 
  pluck("ratioOfPupils9thGradeEligibleForNationalProgramNATE")

```

So that works to get one list out. Now let's do all that contain a variable named `value` and get them in a single dataframe. A relatively simple way to do this (in the current data) is to filter the list elements that contain more than one value.

```{r}
vars <- which(sapply(test$body, function(x) length(x) > 1))
names(vars)
```

Then we can bind them together.

```{r}

# create empty dataframe to store output of loop in
df_total <- data.frame()

for (i in names(vars)){
  tmp <- test$body %>% 
    pluck(i) %>% 
    add_column(variable = i)
  df_total <- rbind(df_total,tmp)
}

glimpse(df_total)
```

Looks good, although there will probably be a lot of recoding needed later.

### Getting data from multiple schools

Since the API demands that we specify the type of school in the API call, we need to add this information to the list of schools. This means that we first need to retrieve the basic information for each school.

```{r}
sdataV3 <- GET("https://api.skolverket.se/planned-educations/v3/school-units/84411355") %>% 
  content("text") %>% 
  fromJSON()

sdataV3$body$typeOfSchooling$code
```

So that is where we find the type code for each school.

#### Check if all schools have data in the database.

```{r}
schoolsAvailable <- data.frame()
for (i in schools$Skolenhetskod) {
  tmp <- http_status(GET(paste0("https://api.skolverket.se/planned-educations/v3/school-units/", i))) %>%
    pluck("reason") %>%
    as.data.frame(nm = "status") %>%
    add_column(Skolenhetskod = i)
  schoolsAvailable <- rbind(schoolsAvailable, tmp)
}

schoolsAvailable %>% 
  count(status)

```

14 schools are not in the database and need to be removed from the list of schools before we retrieve data. Maybe they match up with the `Status` variable?
```{r}
schools %>% 
  left_join(.,schoolsAvailable, by = "Skolenhetskod") %>% 
  filter(!status == "OK")
```
Indeed, but not a perfect match, since the numbers don't add up when compared to this:

```{r}
schools %>% 
  count(Status)
```

Proceeding to remove schools unavailable in database.

```{r}

schoolsFiltered <- schools %>% 
  left_join(.,schoolsAvailable, by = "Skolenhetskod") %>% 
  filter(status == "OK") %>% 
  select(!status)

schoolTypes <- data.frame()

for (i in schoolsFiltered$Skolenhetskod) {
  
  tmp <- GET(paste0("https://api.skolverket.se/planned-educations/v3/school-units/", i)) %>% 
    content("text") %>% 
    fromJSON()
  tmp2 <- data.frame(
    type = tmp$body$typeOfSchooling$code,
    Skolenhetskod = i)
                     
  schoolTypes <- rbind(schoolTypes,tmp2)
}
  
glimpse(schoolTypes)
```

Hmm. We have 82 schools now, rather than 52. Maybe some schools have multiple types?

```{r}
schoolTypes %>% 
  count(Skolenhetskod) %>% 
  filter(n > 1)
```

#### Data retrieval

Since we have two variables to loop over, we could use a nested `for()` loop, but we could also use `map2()` to retrieve data for all schools.

```{r}
schoolData <- map2(
  .x = schoolTypes$Skolenhetskod,
  .y = schoolTypes$type,
  ~ GET(paste0("https://api.skolverket.se/planned-educations/v3/school-units/", .x,"/statistics/", .y)) %>%
    content("text") %>%
    fromJSON()
)

```

### Unnesting multiple lists

Next step is to get each schools data from a list element to a dataframe, and then combine all of them. We already did the first part for one school, so let's expand on that.

We'll define a function to get the data from one school.

```{r}
oneSchool <- function(listN) {
  schoolUnit <- schoolData[[listN]]$body$schoolUnit
  df_total <- data.frame()

  for (i in names(vars)) {
    tmp <- schoolData[[listN]]$body %>%
      pluck(i) %>%
      add_column(variable = i,
                 Skolenhetskod = schoolUnit)
    df_total <- rbind(df_total, tmp)
  }
  return(df_total)
}
```

Test the function.

```{r}
oneSchool(1) %>% 
  head(10)
```

Looks good.

```{r}
#| eval: false
nestedOutput <- map(c(1:length(schoolData)), ~ oneSchool(.x))
```

This fails at index 2. **(Code is not run since it won't allow the output to be rendered)**

### Troubleshooting

```{r}
schoolData[[2]]$body
```

Looks like not all schools have the same data variables. We'll have to work that into the function

```{r}
oneSchool <- function(listN) {
  schoolUnit <- schoolData[[listN]]$body$schoolUnit
  vars <- which(sapply(schoolData[[listN]]$body, function(x) length(x) > 1))
  df_total <- data.frame()

  for (i in names(vars)) {
    tmp <- schoolData[[listN]]$body %>%
      pluck(i) %>%
      add_column(variable = i,
                 Skolenhetskod = schoolUnit)
    df_total <- rbind(df_total, tmp)
  }
  return(df_total)
}
```


```{r}
#| eval: false
nestedOutput <- map(c(1:length(schoolData)), ~ oneSchool(.x))

```

Failing at index 6... **(Code is not run since it won't allow the output to be rendered)**

```{r}
schoolData[[6]]$body$programMetrics %>% 
  head(5)
```

There is a variable that has more than 3 columns.

```{r}
schoolData[[6]]$body$schoolUnit
```

No unit code in the data.

```{r}
schoolTypes %>% 
  slice(6)
```

This is a Gymnasieskola.

```{r}
schoolTypes %>% 
  rownames_to_column() %>% 
  filter(type == "gy")
```

Looks like we have 5 of those. Maybe they all share the same structure?

```{r}
map(c(6,9,10,16,74), ~ names(schoolData[[.x]]$body))
```

Indeed. We'll just filter those schools out for now.

```{r}
mapSchoolUnits <- schoolTypes %>% 
  rownames_to_column() %>% 
  filter(!type == "gy") %>% 
  pull(rowname) %>% 
  as.numeric()
```


```{r}
#| eval: false
nestedOutput <- map(mapSchoolUnits, ~ oneSchool(.x))

```

Index 19 error.

```{r}
schoolData[[mapSchoolUnits[19]]]
```

Strange that this school has no data when we tested that earlier. But alas, we'll just have to remove it. Note that the type is "vux". We should have filtered on types of interest earlier. This is not really a relevant type for our purposes.

```{r}
schoolTypes %>% 
  count(type)
```

Let's get the full descriptions of these abbreviations.

```{r}
schoolTypes <- data.frame()

for (i in schoolsFiltered$Skolenhetskod) {
  
  tmp <- GET(paste0("https://api.skolverket.se/planned-educations/v3/school-units/", i)) %>% 
    content("text") %>% 
    fromJSON()
  tmp2 <- data.frame(
    type = tmp$body$typeOfSchooling$code,
    typeDesc = tmp$body$typeOfSchooling$displayName,
    Skolenhetskod = i)
                     
  schoolTypes <- rbind(schoolTypes,tmp2)
}

schoolTypes %>% 
  count(typeDesc,type)
```

For our purposes, we are primarily interested in gr and gy, and some stats from fsk, but not the rest. This filtering should of course have been done earlier, but the first part of this post was written on a Saturday after 2 weeks of vacation and my brain was not quite up to speed...

### Data retrieval with filtering on school types

We'll do them separately this time, starting with "gr" only.

```{r}
mapSchoolUnits <- schoolTypes %>% 
  rownames_to_column() %>% 
  filter(type == "gr") %>% 
  pull(rowname) %>% 
  as.numeric()
         
nestedOutputGR <- map(mapSchoolUnits, ~ oneSchool(.x))

glimpse(nestedOutputGR[[1]])
```

Now each school has its own dataframe, stored within the list object `nestedOutputGR`. Next step is to combine them into one dataframe.

### Combining dataframes

```{r}
unnestedOutputGR <- map_dfr(c(1:length(nestedOutputGR)), ~ do.call(bind_rows, nestedOutputGR[[.x]]))
glimpse(unnestedOutputGR)
```

Data is now in long format, which will be useful for some uses, and we can use `pivot_wider()` to change format when needed.

All variables are class "character", which is fine for some, but `timePeriod` is a date variable, and `value` has "." for missing and "," as decimal sign. This needs to be modified.

## Variables

Let's have a look at the variables available in the data.

```{r}
unnestedOutputGR %>% 
  count(variable)
```

### Types

```{r}
unnestedOutputGR %>% 
  count(valueType)
```

Let's look at the value variable for the different valueTypes (except EXISTS).

```{r}
unnestedOutputGR %>% 
  filter(!valueType == "EXISTS") %>% 
  group_by(valueType) %>% 
  distinct(value)
```

We need the value variable to be of class numeric, which means only numerics and NA values are allowed.

### Recoding

```{r}
# create copy of dataframe prior to recoding
df.gr <- unnestedOutputGR

# recode . to NA
df.gr$value <- car::recode(df.gr$value,"'.'=NA")
# change decimal comma to decimal point
df.gr$value <- gsub(pattern = ",",
                    replacement = ".",
                    df.gr$value)
```


```{r}
# remaining cleanup
df.gr$value <- car::recode(df.gr$value,"'..'=NA") # too few pupils
df.gr$value <- car::recode(df.gr$value,"'*'=NA") # no req for teacher license
df.gr$value <- gsub("cirka ", "", df.gr$value)
df.gr$value <- gsub("~", "", df.gr$value)

df.gr$value <- as.numeric(df.gr$value)
```

#### Date variable

First remove the second part of the school year designation (extract the first 4 digits).
```{r}
# date variable
df.gr$timePeriod <- str_extract(df.gr$timePeriod, "\\d{4}")
df.gr$timePeriod <- as.integer(df.gr$timePeriod)

df.gr %>% 
  count(timePeriod)
```

We could transform this to a date format variable, but having dates as year numerics is fine for our intended use.

## Visualizing data

```{r}
# we like to use school names instead of codes when creating figures
df.gr <- left_join(df.gr,schools, by = "Skolenhetskod")
```


### Variable check

```{r}
library(gtsummary)

df.gr %>% 
  select(timePeriod,variable) %>% 
  group_by(variable) %>% 
  tbl_summary(timePeriod)

```

### Number of pupils per school

```{r}
df.gr %>% 
  filter(variable == "totalNumberOfPupils") %>% 
  ggplot(aes(x = factor(timePeriod), 
             y = value, 
             color = Skolenhetsnamn,
             group = Skolenhetsnamn)) +
  geom_point() +
  geom_line() +
  theme_bw()
```

Only data from 2022 for that variable. We need to look elsewhere for historical data, it seems. It would also be useful to have the number of pupils per class, since we want to make comparisons to survey data responses. However, the documentation of this variable indicates that number of pupils is rounded to nearest 10th:

> "Antal elever på skolenheten, avrundat till tiotal.  1-14 elever ger 10, 15-24 elever ger 20, 25-34 elever ger 30, enkelprick förekommer om elever saknas.
Uppgiften samlas in 15 oktober varje år och uppdatering sker i mars året efter insamling."

Also, we can see that there are several school units that have very similar names, or represent parts of the same schools. This data is clearly quite complex in its structure.

As a bonus item, here is a more suitable type of figure to illustrate data from a single year, 

```{r}
df.gr %>% 
  filter(variable == "totalNumberOfPupils") %>% 
  mutate(Skolenhetsnamn = factor(Skolenhetsnamn)) %>% 
  ggplot(aes(x = fct_reorder(Skolenhetsnamn, desc(value)), 
             y = value,
             fill = value)) +
  geom_col() +
  theme_minimal() +
  labs(x = "",
       y = "Number of pupils") +
  scale_fill_viridis_c(guide = "none") +
  coord_flip()

```

### studentsPerTeacherQuota

Here we have data from multiple school years.

```{r}
df.gr %>% 
  filter(variable == "studentsPerTeacherQuota") %>% 
  ggplot(aes(x = factor(timePeriod), 
             y = value, 
             color = Skolenhetsnamn,
             group = Skolenhetsnamn)) +
  geom_point() +
  geom_line() +
  theme_bw()
```

Huge variation, probably since there are schools in the data that seem to represent special education groups. Let's look at the average numbers across years and sort out the 10 with the lowest number of pupils per teacher.

```{r}
df.gr %>% 
  filter(variable == "studentsPerTeacherQuota") %>% 
  group_by(Skolenhetsnamn) %>% 
  reframe(Average = mean(value)) %>% 
  arrange(Average) %>% 
  slice(1:10)
```

We'll remove the ones with "1GR" at the end of the school name.

```{r}
df.gr %>% 
  filter(variable == "studentsPerTeacherQuota",
         !str_detect(Skolenhetsnamn,"1GR")) %>% 
  ggplot(aes(x = factor(timePeriod), 
             y = value, 
             color = Skolenhetsnamn
             )) +
  geom_point() +
  geom_line(aes(group = Skolenhetsnamn)) +
  theme_bw()
```

Let's update our information about the schools with some more variables.

```{r}

schoolUnitCodes <- schoolTypes %>% 
  filter(type == "gr") %>% 
  pull(Skolenhetskod)

schoolTypes2 <- data.frame()

for (i in schoolUnitCodes) {
  
  tmp <- GET(paste0("https://api.skolverket.se/planned-educations/v3/school-units/", i)) %>% 
    content("text") %>% 
    fromJSON()
  
  if (length(tmp$body$typeOfSchooling$code) > 1){ # if both fsk and gr
  
    tmp2 <- data.frame(
    type = tmp$body$typeOfSchooling$code[2],
    principalOrganizerType = tmp$body$principalOrganizerType,
    companyForm = tmp$body$companyForm,
    schoolYears = paste0(tmp$body$typeOfSchooling$schoolYears[2][[1]][1],"-",tmp$body$typeOfSchooling$schoolYears[2][[1]][length(tmp$body$typeOfSchooling$schoolYears[2][[1]])]),
    corporationName = tmp$body$corporationName,
    latitude = tmp$body$wgs84_Lat,
    longitude = tmp$body$wgs84_Long,
    zipCode = tmp$body$contactInfo$addresses$zipCode[2],
    Skolenhetskod = i)
                     
  schoolTypes2 <- bind_rows(schoolTypes2,tmp2)
  } else { # when schoolType only "gr"
    tmp2 <- data.frame(
    type = tmp$body$typeOfSchooling$code,
    principalOrganizerType = tmp$body$principalOrganizerType,
    companyForm = tmp$body$companyForm,
    schoolYears = paste0(tmp$body$typeOfSchooling$schoolYears[1][[1]][1],"-",tmp$body$typeOfSchooling$schoolYears[1][[1]][length(tmp$body$typeOfSchooling$schoolYears[1][[1]])]),
    corporationName = tmp$body$corporationName,
    latitude = tmp$body$wgs84_Lat,
    longitude = tmp$body$wgs84_Long,
    zipCode = tmp$body$contactInfo$addresses$zipCode,
    Skolenhetskod = i)
                     
  schoolTypes2 <- bind_rows(schoolTypes2,tmp2)
    }
}

schoolTypes2 <- left_join(schoolTypes2,schools, by = "Skolenhetskod")

glimpse(schoolTypes2)
```

Well, that was a handful to work out. And we only get "gr" schools, so a similar process will be necessary if we want to get the other school types.

